Parallel Tractability of Ontology\\ Materialization: Technique and Practice

### 2018-01-27 parallel-tractability-J-v9/response.tex

Indeed, the abstract has been shortened significantly now.

\section{Rebuttal to Review 2}

\begin{quote}
First, in the discussion under Lemma 1, it says $|P^*|$ is polynomial in the size of $P$, which is imprecise, as $|P^*|$ is polynomial in the size of $I$ but not necessarily in the size of $R$. This is not a big issue if one considers NC for data complexity. What is less straightforward and more of concern is the statement that Step~2 costs only one time unit. As Step~2 involves checking the applicability of $B_1, \ldots, B_n \to H$ in $G$ and updating $G$, it depends on the size of $G$ and the size of $G$ is not a constant w.r.t.\ that of $I$. Footnote~4 suggests checking applicability of the rule can be done in constant time via an index. What about updating $G$ and maintaining the index?
\end{quote}

In the paragraph under Lemma~1, we made it clear that $|P^*|$ is polynomial in the size of $I$ for a class of datalog programs.

The statement that Step~2 costs only one time unit is based on the context of one processor. Since one processor
is allocated one rule instance (see the first paragraph of Algorithm~$\mathsf{A}_{bsc}$), 
the procedure can be completed in constant time units. We rewrote these sentences by clarifying this.

(TODO) We added a paragraph to discuss the complexity of Algorithm~$\mathsf{A}_{bsc}$.
The constant time of checking applicability ......

@@

\begin{quote}
The computation of the rch relation $S_{\text{rch}}$ ($\dag$) looks problematic, as it requires ``H has been added to G''. If it were the case, then in Example~4 the $S_{\text{rch}}$ would always be empty, as none of the rule heads $H$ would be added to $G$. Again, the complexity analysis of $A_{\text{opt}}$ is unclear and possibly flawed. Step~2 uses an NC algorithm to compute the transitive closure of $S_{\text{rch}}$, which runs in poly-logarithmic time w.r.t.\ the size of $S_{\text{rch}}$ not necessarily so w.r.t.\ the size of $I$. What is the size of $S_{\text{rch}}$ w.r.t.\ that of $I$? Similarly as above, it involves updating $G$, which depends on the sizes of both $G$ and $S_{\text{rch}}$. Would it be still in poly-logarithmic time?
\end{quote}

The requirement ``H has been added to G'' in the description of computing the rch relation $S_{\text{rch}}$ ($\dag$) is indeed
incorrect. It should be ``H has not been added to G''. We fixed this error.

(TODO) We added a paragraph to discuss the complexity of Algorithm~$\mathsf{A}_{opt}$.
In Step~2 of Algorithm~$\mathsf{A}_{opt}$, an NC algorithm is used to compute the transitive closure of $S_{\text{rch}}$.
It can be checked that the scale of $S_{\text{rch}}$ is less than that of $P^*$ and is in polynomial of
the size of inputs. Thus, the NC complexity of the computation of the transitive closure can be guaranteed.

@@

\begin{quote}
Finally, the algorithms assume each processor handling a rule instantiation. I wonder how it is implemented in ParallelDHL.
\end{quote}

(TODO) We used a paragraph to discuss this after Algorithm~$\mathsf{A}_{prc}$.

@@

\begin{quote}
I list some relatively minor issues regarding presentation and technical details:

- p20, it mentions a group of materialisation rules in YAGO which belong to $D_{\text{dhl}}$. It deserves a clarification on what these rules are like and why they do not fall into existing parallel tractable classes.\\
- p20, Table 2, it should mention how these ontologies were selected. \\
- p21, the discussion on line graph lg1 is confusing (i.e.\ ``line-2 stays higher than line-6, but lower than line-4''). Maybe I was looking at the wrong place. \\
- p25, the section heading ``Reference'' occur twice.
\end{quote}

### 2018-01-27 parallel-tractability-J-v9/parallel-tractability-J.tex
fixed some quotes of reviewer 2

We now discuss how Algorithm~$\mathsf{A}_{bsc}$ can be restricted to an \texttt{NC} version.
(\uppercase\expandafter{\romannumeral1}) Since Algorithm~$\mathsf{A}_{bsc}$ does not introduce
new constants and each predicate has a constant arity,
one can check that $|P^*|$ is polynomial in the size of $P$. This also means that
the number of processors is polynomially bounded.
(\uppercase\expandafter{\romannumeral2}) The computing time of \ref{alg1:addFacts}
and \ref{alg1:halt} occupies constant time units because of parallelism.
(\uppercase\expandafter{\romannumeral3}) The main computation part in Algorithm~$\mathsf{A}_{bsc}$ is the
iteration of \ref{alg1:updateG}.
In each iteration of \ref{alg1:updateG}, all processors work independently from each other.
Thus, in theory, \ref{alg1:updateG} costs one time unit.
The whole computing time turns out to be bounded by the number of iterations of \ref{alg1:updateG}. (\uppercase\expandafter{\romannumeral4})
We use the symbol $\psi$ to denote a poly-logarithmically bounded function.
The input of $\psi$ is the size of $P$ and the output is a non-negative integer.
Based on (\uppercase\expandafter{\romannumeral1}, \uppercase\expandafter{\romannumeral2},
\uppercase\expandafter{\romannumeral3}, \uppercase\expandafter{\romannumeral4}),
for any datalog program $P$, if we use $\psi(|P|)$ to
bound the number of iterations of \ref{alg1:updateG}, then Algorithm~$\mathsf{A}_{bsc}$
is an \texttt{NC} algorithm, denoted by $\mathsf{A}_{bsc}^{\psi}$ (this symbol also
means that $\psi$ occurs as a configuration for Algorithm~$\mathsf{A}_{bsc}$).

@@

Consider Example~\ref{exp:mg} again. Let the integer $k$ be a variable. We can
get a class of datalog programs, denoted by $\mathbb{P}_{ex_1}$, where the rule set
is $\{R(x,y),A(y)\rightarrow A(x), S(x,y),R(y,z)\rightarrow R(x,z)\}$ and the instance
set varies according to $k$. The algorithm $\mathsf{A}_{bsc}^{\psi}$ is restricted in the sense that it cannot even work on the rather simple datalog program class $\mathbb{P}_{ex_1}$.
It can be checked that, for some materialization graph $\mathcal{G}_{ex_1}$
corresponding to a datalog program in $\mathbb{P}_{ex_1}$, \texttt{depth}($\mathcal{G}_{ex_1}$)$=k$ for some $k$.
This means that the depths of the materialization graphs are linearly bounded by $k$.
On the other hand, the sizes of the datalog programs in $\mathbb{P}_{ex_1}$ are polynomial of $k$.
Thus, for any $\psi$ that is poly-logarithmically bounded, we can always find a $k$ large
enough such that $\mathsf{A}_{bsc}^{\psi}$ terminates without constructing a materialization
graph for each datalog program in $\mathbb{P}_{ex_1}$.
However there indeed exists an \texttt{NC} algorithm that can handle $\mathbb{P}_{ex_1}$.
We discuss this in the next part.

@@

\begin{description}[leftmargin=2ex]
\item[(\textbf{\dag})] \emph{For each rule instantiation of the form $B_1,..,B_i,..,B_n\rightarrow H$
where $H$ has been added to $\mathcal{G}$}:
\begin{enumerate}[leftmargin=2ex]
\item \emph{if the body atoms $B_1,...,B_n$ have all been added to $\mathcal{G}$, put \texttt{rch}$(B_1,H),...,$ \texttt{rch}$(B_n,H)$ in $S_{\textit{\tiny rch}}$};
\item \emph{if $B_i$ is the only node in the body that has not been added to $\mathcal{G}$,
    put \texttt{rch}$(B_i,H)$ in $S_{\textit{\tiny rch}}$}.\hfill$\Box$
\end{enumerate}
\end{description}

>>>>

We now discuss how Algorithm~$\mathsf{A}_{bsc}$ can be restricted to an \texttt{NC} version.
(\uppercase\expandafter{\romannumeral1}) Given a class of datalog programs $\mathbb{P}$ where
a rule set is shared for each datalog program $P=\langle R, \textbf{I}\rangle$ in $\mathbb{P}$,
it can be checked that $|P^*|$ is polynomial in the size of $\textbf{I}$. 
This is because that each predicate occurring in $\mathbb{P}$ has a constant arity
and Algorithm~$\mathsf{A}_{bsc}$ does not introduce
new constants. This also means that the number of processors is polynomially bounded.
(\uppercase\expandafter{\romannumeral2}) The computing time of \ref{alg1:addFacts}
and \ref{alg1:halt} occupies constant time units because of parallelism.
(\uppercase\expandafter{\romannumeral3}) The main computation part in Algorithm~$\mathsf{A}_{bsc}$ is the
iteration of \ref{alg1:updateG}. Since each processor handles one rule instance, 
\ref{alg1:updateG} costs one time unit on each processor in theory. 
Assume that there is no delay and other communication problems among all the processors.
We have that the computing time of \ref{alg1:updateG} of the whole algorithm costs one time unit as well.
Further, the whole computing time turns out to be bounded by the number of iterations of \ref{alg1:updateG}. (\uppercase\expandafter{\romannumeral4})
We use the symbol $\psi$ to denote a poly-logarithmically bounded function.
The input of $\psi$ is the size of $P$ and the output is a non-negative integer.
Based on (\uppercase\expandafter{\romannumeral1}, \uppercase\expandafter{\romannumeral2},
\uppercase\expandafter{\romannumeral3}, \uppercase\expandafter{\romannumeral4}),
for any datalog program $P$, if we use $\psi(|P|)$ to
bound the number of iterations of \ref{alg1:updateG}, then Algorithm~$\mathsf{A}_{bsc}$
is an \texttt{NC} algorithm, denoted by $\mathsf{A}_{bsc}^{\psi}$ (this symbol also
means that $\psi$ occurs as a configuration for Algorithm~$\mathsf{A}_{bsc}$).

@@

Consider Example~\ref{exp:mg} again. Let the integer $k$ be a variable. We can
get a class of datalog programs, denoted by $\mathbb{P}_{ex_1}$, where the rule set
is $\{R(x,y),A(y)\rightarrow A(x), S(x,y),R(y,z)\rightarrow R(x,z)\}$ and the fact
set varies according to $k$. The algorithm $\mathsf{A}_{bsc}^{\psi}$ is restricted in the sense that it cannot even work on the rather simple datalog program class $\mathbb{P}_{ex_1}$.
It can be checked that, for some materialization graph $\mathcal{G}_{ex_1}$
corresponding to a datalog program in $\mathbb{P}_{ex_1}$, \texttt{depth}($\mathcal{G}_{ex_1}$)$=k$ for some $k$.
This means that the depths of the materialization graphs are linearly bounded by $k$.
On the other hand, the sizes of the datalog programs in $\mathbb{P}_{ex_1}$ are polynomial of $k$.
Thus, for any $\psi$ that is poly-logarithmically bounded, we can always find a $k$ large
enough such that $\mathsf{A}_{bsc}^{\psi}$ terminates without constructing a materialization
graph for each datalog program in $\mathbb{P}_{ex_1}$.
However there indeed exists an \texttt{NC} algorithm that can handle $\mathbb{P}_{ex_1}$.
We discuss this in the next part.

@@

\begin{description}[leftmargin=2ex]
\item[(\textbf{\dag})] \emph{For each rule instantiation of the form $B_1,..,B_i,..,B_n\rightarrow H$
where $H$ has not been added to $\mathcal{G}$}:
\begin{enumerate}[leftmargin=2ex]
\item \emph{if the body atoms $B_1,...,B_n$ have all been added to $\mathcal{G}$, put \texttt{rch}$(B_1,H),...,$ \texttt{rch}$(B_n,H)$ in $S_{\textit{\tiny rch}}$};
\item \emph{if $B_i$ is the only node in the body that has not been added to $\mathcal{G}$,
    put \texttt{rch}$(B_i,H)$ in $S_{\textit{\tiny rch}}$}.\hfill$\Box$
\end{enumerate}
\end{description}
